% Chapter 7

\chapter{Conclusion And Future Work} % Main chapter title

\label{Chapter7} % For referencing the chapter elsewhere, use \ref{Chapter6} 

\lhead{Chapter 7. \emph{Conclusion And Future Work}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------


\section{Conclusion}
In chapter 1, the project's main idea was illustrated which
is a web service that provides automatic email categorization into
user-defined folders based on machine learning and data mining
classification techniques.

A classification web service was built with REST API \cite{REST} that receives
a classification request with the required email to be classified and returns
the most suitable category for this email.

An Email monitoring web service was built as a client for the classification
web service mentioned above. Users can register in this web service to have
their email accounts continuously monitored and their incoming emails automatically
classified to the best category.

A Google Chrome Browser \cite{CHROME} Extension was built as another example of a
classification web service client. The Extension provides google chrome users
with a graphical user interface that is integrated with gmail web interface.
Using this extension , Users can send email classification requests on demand.
The extension also detects manual classification done by the user to be sent
for the classification web service as a feedback on the classifier to learn
from the users classification pattern.
%================================================
\section{Future Work}
In the previous section, the implemented features were stated. In this section
the new features that could be added to Smart Email will be introduced.

\subsection{Future work related to classification web service}
\begin{itemize}
    \item Security issues: Emails privacy and security are considered very
    important issues, so an additional security to communication between clients and
    web service need to supported.
    \item Multi-Label email classification: Currently the classification web
    service assigns each email only one label. Support for multi-label
    classification can be added.
    \item Improving the classification accuracy.
\end{itemize}

\subsection{Future work related to other Smart Email components}
\begin{itemize}
    \item summarization of email threads can be supported.
    \item Supporting other browsers like (Firefox and Internet Explorer).
\end{itemize}


