% Chapter 1

\chapter{Introduction} % Main chapter title

\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1} 

\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Motivation}

The invention of email permanently changed the way communication takes place. 
Email messages provide fast, free and reliable method for communication. 
That's why email has become the most frequently used method for communication 
nowadays.

With the increasing popularity of emails, users spend a significant portion of 
their time reading and replying to emails in their inbox folder. One solution 
is to organize emails into several folders -one folder for each topic- rather 
than one big congested folder which is the inbox. This facilitates the process 
of managing, prioritizing and facilitating the search process for emails. 
The problem is that the process of email categorization has to be done manually, 
which tends to be tedious and time consuming for most users.

Several machine learning techniques have been applied for solving different 
email related problems such as spam detection. In our work, the problem of 
automatic email categorization into folders is addressed, that is, assigning 
email messages into user-specific folders automatically. This task provides 
a valuable addition to email clients by avoiding mailboxes cluttering and 
providing a better automated way for managing and structuring incoming email 
messages.

Email classification introduces new challenges not found in ordinary text 
classification. This is related to the special nature of email messages. Email 
messages tend to be small in size compared to ordinary text documents. User 
created folders don't necessarily correspond to a single semantic topic, 
folders might contain email messages for unfinished todos, or emails based on 
a certain sender or recipient. Email messages also arrive in a stream over time 
which causes more difficulties because the main topic may drift over time.

Smart Email is created to address these problems. Smart Email is trained on the 
existing user folders and tries to automatically classify new incoming emails 
into the best folder they could belong to. Smart Email works as a classification
 web service that receives classification requests and responds with the best
folder for the email. The email monitoring web service allows each user to
 register with one or more email account and provides the user with a central place
 to monitor and control the classification procedure for his emails.

\section{Scope of Work}

The problem of automatic email categorization into folders was introduced
in the previous section. Smart Email is an application that provides a solution
for this problem.

Users receive new emails every day. The email monitoring web service monitors
 the user's account and triggers the classification procedure on any new email. 

The Smart Email classification service receives the classification request with the new email
 and performs the classification procedure to determine the best category for this new 
email. The new Email is moved to the predicted category folder, when the user 
opens his email account he will notice the new changes right away. No raw emails 
are stored at the server side after the classification procedure has taken place.

The user provides feedback on the performance of the classification so that the 
classifier can re-learn from the misclassified emails. The classification service maintains 
a classification model for each user which can be updated and improved incrementally 
based on the user feedback.

The email monitoring web service enables the user to register with multiple email accounts and 
provides him with a central view to monitor and control everything related to 
the email classification procedure.


\section{Organization of the Report}

The report is organized into 7 chapters. In chapter \ref{Chapter2}, a background 
about email classification techniques will be discussed, and related work and applications
which tried to solve the problem will be presented in sections \ref{sec:related_work} 
and \ref{sec:related_apps}. In chapter \ref{Chapter3}, suggested features of Smart 
Email Classifier will be illustrated.
Smart email requirements and design will be introduced in chapter \ref{Chapter4}.
Chapter \ref{Chapter5} will include development process, environment and tools. 
Implementation details, results and discussion will be explained in chapter \ref{Chapter6}. Finally,  
chapter \ref{Chapter7} will contain the conclusion and future work of Smart Email.


